Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die Einhaltung der Echtzeit

### 2024-05-06 Masterarbeit_Robotik_Pamuk.tex
Organize content

\chapter{Salamander 4}\label{cha:salamander4}

@@

Xenomai consists of 3 parts. These can be found in the Table \ref{tab:what_is_xenomai}.

@@

\clearpage
\section{Salamander 4 Description}

\subsection{Task priorities}
Es gibt grundsätzlich 4 Gruppen

@@

\subsection{Memory Management}
Es gibt verschiedene Speicherbereiche

@@

Eine LASAL CPU besteht aus den folgenden Software-Modulen:
\begin{itemize}
	\item Operating system
	\item Loader
	\item Hardware-Klassen
\end{itemize}

Die Schnittstelle zwischen den einzelnen Modulen wird in Abbildung \ref{fig:lasal_cpu} durch einen Pfeil gekennzeichnet.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
	\caption[LASAL CPU]{LASAL CPU}
	\label{fig:lasal_cpu}
\end{figure}

@@

\section{Salamander 4 Bare Metal}
Salamander 4 Bare Metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system, including

@@

Figure \ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\columnwidth]{img/max_latency_vapic.png}
	\caption[Latency taskset]{Latency taskset}
	\label{fig:max_latency_vapic}
\end{figure}

@@

\section{Latency Comparison}
In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4. For this purpose, the latency tool of the Xenomai test suite was used. The latency was measured under two conditions, idle and CPU-stressed. The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.

@@

After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
\clearpage

@@

@@

\section{KVM exit reasons}\label{sec:kvm_exit_reasons}

\clearpage
\subsection{HLT}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{EPT\_MISCONFIG}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{PREEMPTION\_TIMER}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{EXTERNAL\_INTERRUPT}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{IO\_INSTRUCTION}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{EOI\_INDUCED}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{EPT\_VIOLATION}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{PAUSE\_INSTRUCTION}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{CPUID}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{MSR\_READ}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage

@@

>>>>

\chapter{Salamander 4}\label{cha:salamander4}
This chapter breifly describes the Salamander 4 operating system by SIGMATEK.

@@

\noindent Salamander 4 is powered by SIGMATEK's CPU and is comprised of the following software modules:

 \begin{itemize}
	\item \textbf{Operating system}
	\item \textbf{Loader}
	\item \textbf{Hardware classes}
	\item \textbf{Application}
\end{itemize}

\noindent The interfaces between the individual modules are indicated by an arrow in Figure \ref{fig:lasal_cpu}.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
	\caption[LASAL CPU]{LASAL CPU}
	\label{fig:lasal_cpu}
\end{figure}
\clearpage
\section{Xenomai}

\noindent Xenomai consists of 3 parts. These can be found in the Table \ref{tab:what_is_xenomai}.

@@

\section{Task priorities}
Es gibt grundsätzlich 4 Gruppen

@@

\section{Memory Management}
Es gibt verschiedene Speicherbereiche

@@

@@

\chapter{Initial Real-Time Latency}\label{cha:initial-real-time-latency}

\clearpage
\section{Salamander 4 Bare Metal}
Salamander 4 Bare Metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system, including

@@

Figure \ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\columnwidth]{img/max_latency_vapic.png}
	\caption[Latency vapic]{Latency vapic}
	\label{fig:max_latency_vapic}
\end{figure}

@@

@@

After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.

\clearpage
\section{Latency Comparison}
In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4. For this purpose, the latency tool of the Xenomai test suite was used. The latency was measured under two conditions, idle and CPU-stressed. The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.

\clearpage
\section{KVM exit reasons}\label{sec:kvm_exit_reasons}

\clearpage
\subsection{HLT}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{EPT\_MISCONFIG}
\subsubsection{Description}
\subsubsection{Occurence Reduction}
\clearpage
\subsection{PREEMPTION\_TIMER}
\clearpage
\subsection{EXTERNAL\_INTERRUPT}
\clearpage
\subsection{IO\_INSTRUCTION}
\clearpage
\subsection{EOI\_INDUCED}
\clearpage
\subsection{EPT\_VIOLATION}
\clearpage
\subsection{PAUSE\_INSTRUCTION}
\clearpage
\subsection{CPUID}
\clearpage
\subsection{MSR\_READ}
\clearpage

@@

\subsection{Tasks and Events}

@@

\vspace{1em}
\begin{minipage}{\linewidth}
	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualisation},label={script:qemu_configuration_optimizations}]
    #!/bin/sh

    exec  qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
    -m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
    -append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable" \
    -net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
    -fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
    -device vhost-vsock-pci,guest-cid=3,id=vsock0 \
    -drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
    -no-reboot -nographic
\end{lstlisting}
\end{minipage}

@@

\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
\cite{Tracecmd}
\cite{KernelShark}
\cite{XenomaiXenomai}
\cite{WelcomeYoctoProject}
\cite{QEMU}
\cite{linPerformanceEvaluationXenomai}
\cite{abediRTNFPredictableLatency2019}
\cite{broskyShieldedProcessorsGuaranteeing2003}
\cite{casiniLatencyAnalysisVirtualization2021}
\cite{cinqueVirtualizingMixedCriticalitySystems2022}
\cite{garcia-vallsChallengesRealtimeVirtualization2014}
\cite{guStateoftheArtSurveyRealTime2012}
\cite{HardRealTime}
\cite{javierperezHowRealTime2022}
\cite{kirovaImpactModernVirtualization2019}
\cite{kiszkaLinuxRealTimeHypervisor}
\cite{lutsykPipelinedMulticoreMachine2020}
\cite{malallahComprehensiveStudyKernel2021}
\cite{masrurVMBasedRealTimeServices2010}
\cite{mckenneyRealTimeVs}
\cite{pielAsymmetricSchedulingLoad2006}
\cite{queirozTestingLimitsGeneralpurpose2023}
\cite{reghenzaniRealTimeLinuxKernel2020}
\cite{yoonRealTimePerformanceAnalysis}
\clearpage
